\section{Hyperproperties of MPC Passive Security}
\label{section-hyperprop}

As discussed in Section \ref{section-related-work} previous authors
have noted connections between the traditional formulation of the MPC
security model- aka the simulator security model- and
hyperproperty-style formulations of program trace behavior
\cite{8429300,wysstar} including probabilistic noninterference. In
this Section we show how to formulate security properties in our
language model by direct appeal to program distributions. In
particular, \emph{noninterference modulo output ($\NIMO$)} is shown to
be sound for MPC passive security. In Section \ref{section-automation}
we will show how $\NIMO$ can be automatically enforced directly via
whole-program analysis, or by automatically verifying composable
properties of circuit components.

\subsection{Passive Security and the Real/Ideal Model}
\label{section-hyperprop-passive}
Here we briefly summarize the formal definition of MPC security. For a
more detailed discussion, the reader is referred to \cite{evans2018pragmatic}.
In this paper, we will consider the \emph{passive} aka
semi-honest aka honest-but-curious variant. In this threat
model, protocol participants can be corrupted, but all clients
follow the rules of the protocol- as opposed to the
\emph{active} aka malicious variant where participants
are allowed to break these rules. In the passive model, the
vulnerability is through information leaked to the adversary
through the views of corrupted participants. While the passive
model also assumes that any clients can potentially be
corrupted, it assumes that no more than half will be-
i.e., it assumes an \emph{honest majority}. 

The real/ideal model posits a \emph{simulator} that exists in an ideal
world, whose goal is to reconstruct observable information in a real
world run of a protocol $\prog$ that correctly implements an ideal
functionality $\idealf$- that is, execution of a configuration
$\config{\store}{\prog}$. We will assume that any $\idealf$ is
deterministic- while MPC generally allows for probabilistic ideal
functionalities, this restriction captures a large and important class
of protocols. We will assume that any protocol $\prog$ includes a
public \emph{output view} $\outv$ where the result is published, where
$0$ is the ``public'' client and thus $\outv$ is of the form
$\view{0}{w}$ for some $w$, i.e., is owned by 0.  Note that the
protocol is \emph{correct} iff the resulting memory contains $\outv
\mapsto \idealf(\store_S)$, where $S$ are the secret inputs (honest
and corrupt) in $\prog$. Note that this implies that a correct
protocol publicizes the same output on the same secret inputs given
any choice of random tape.

The simulator is given just the corrupt inputs to the run- that is,
$\store_{S_C}$- and the output of the protocol- that is,
$\idealf(\store)$- and aims to reconstruct the distribution of
adversarial views observed in the real world. The intuition here is
that the simulator is given the information that is necessarily
available to the adversary by observing a run of the protocol, and if
the simulator can reconstruct whatever information appears in the real
world adversarial views, by reconstructing their distributions, then
those views leak no additional information, i.e., they have no
additional dependencies on honest input secrets than what is revealed
by the output and conditionings on corrupt inputs.

The simulator is represented by a probabilistic algorithm $\SIM_C$,
aka a \emph{simulation}, that is parameterized by corrupt inputs and
the output of an ideal functionality, and that returns a set of
adversarial views (as a memory) with some probability. Given
corrupt inputs $\store$ and ideal functionality output $v$,  
we write
$
\prob(\SIM_C(\store,v) = \store')
$
to denote the probability that $\SIM_C(\store,v)$
returns corrupt views $\store'$ as a result. We can then define the
probability distribution of corrupt views reconstructed
by the simulator as follows:
\begin{definition}
  Given $C$, $\store$, and $v$, we write $\progd(\SIM_C(\store,v))$ to
  denote the distribution of corrupt views reconstructed by the
  simulation, where for
  all $\store' \in \mems(V)$:
  $$
  \progd(\SIM_C(\store,v))(\store')\ \defeq\ \prob(\SIM_C(\store,v) = \store') 
  $$
\end{definition}

Then we can define passive security in the real/ideal
model as follows. 
\begin{definition}[Passive Security]
  Assume given a program $\prog$ that correctly implements an ideal
  functionality $\idealf$, with $\iov(\prog) = (S,V)$.  Then $\prog$
  is \emph{passive secure in the simulator model} iff for all
  partitions of the federation into honest and corrupt sets $H$ and $C$
  with $|C| < |H|$ and for all $\store \in \mems(S)$ there exists a
  simulation $\SIM_C$ such that:
  $$
  \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \condd{\progd(\prog)}{V_C}{\store}
  $$
\end{definition}

\subsection{Probabilistic Noninterference ($\pni$)} 
\label{section-hyperprop-ni}
Previous work has noted
relations between probabilistic noninterference and security
properties such as memory trace obliviousness \cite{darais2019language} and passive
security \cite{barthe2019probabilistic}. And above (Section \ref{section-pmf-examples}) we
have demonstrated that our basic encryption scheme enjoys a noninterference
property. To formulate probabilistic noninterference in
our setting we begin by defining low equivalence of memories.
\begin{definition}[Memory Low Equivalence]
  Given $\store^1$ and $\store^2$ with $\dom(\store^1) = \dom(\store^2) = X$,
  we write $\store^1 =_C \store^2$ iff $\store^1_{X_C} = \store^2_{X_C}$.
\end{definition}
Now, we can define probabilistic noninterference as a property of
programs ensuring that low-equivalent initial memories produce
the same low-observable trace (i.e., corrupt views) \emph{with the
same probability}. 
\begin{definition}[Probabilistic Noninterference]
  \label{definition-PNI}
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, we say that
  $\prog$ satisfies \emph{probabilistic noninterference}, and we write
  $\pni(\prog)$, iff for all
  $\store_1, \store_2 \in \mems(S)$:
  $$\store_1 =_C \store_2 \implies
    \condd{\progd(\prog)}{V_C}{\store_1} = \condd{\progd(\prog)}{V_C}{\store_2}$$
\end{definition}

Despite the intuitive resonance between noninterference and passive
security, since MPC protocols implement an ideal functionality, they
cannot generally satisfy noninterference. Noting that the public
output is always a corrupt view, we can easily demonstrate this by
using Example \ref{example-gmw-andcircuit} as a failure witness, where
we'll assume $C = \{0,2\}$ and $H = \{1\}$:
$$
(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 0, \sx{2}{s2} \mapsto 1 \} \})
= 0
$$
whereas:
$$
(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 1, \sx{2}{s2} \mapsto 1 \} \})
= 1
$$

Probabilistic noninterference can also be restated to formally capture a
basic intuition that noninterfering programs do not provide
the adversary with any clues about honest secrets through corrupt
views, as in the following Lemma. This is similar to asserting that
adversarial knowledge is not increased by the execution of a program as in formulations
of so-called gradual release \cite{4223226}. 
\begin{lemma}
  \label{lemma-pni}
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, 
  $\prog$ satisfies probabilistic noninterference iff for all
  $\store \in \mems(S \cup V)$,
  $\condd{\progd(\prog)}{S_H}{\store_{S_C}} =
    \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}}$.
\end{lemma}

\subsection{Noninteference Modulo Output ($\NIMO$)}
\label{section-nimo}
The intuition that passive security means that the protocol releases
no more information than what is released by conditioning on the
public output and corrupt inputs suggests a
refinement of noninterference. That is, the set of low equivalent
initial memories \emph{that can produce the same output} should generate the
same distributions of adversarial views.
\begin{definition}[Noninterference Modulo Output]
  \label{definition-NIMO}
  We say that a program  $\prog$ satisfies \emph{noninterference modulo output},
  written $\NIMO(\prog)$, iff for all $H$ and $C$ with $|C|\le|H|$ and 
  $\store_1,\store_2 \in \mems(S)$ we have:
  $$
  (\store_1 =_C \store_2 \ \wedge \ 
  (\condd{\progd(\prog)}{\{ \outv \}}{\store_1} = \condd{\progd(\prog)}{\{ \outv \}}{\store_2}))
  \implies 
  (\condd{\progd(\prog)}{V_C}{\store_1} = \condd{\progd(\prog)}{V_C}{\store_2})
  $$
  where $\iov(\prog) = (S,V)$.
\end{definition}

From an information flow perspective, we can say that a $\NIMO$
protocol does not release information beyond what is declassified
via the public output. From a real/ideal perspective, we can
observe that $\NIMO$ implies that the simulator can use
the protocol itself to reconstruct the distribution of adversarial
views. That is, given the output of the ideal functionality
$\idealf(\store_S)$, the simulator can randomly choose any
$\store'$ such that $\store' =_C \store$ and $\idealf(\store'_S) =
\idealf(\store_S)$ and run the protocol with initial memory $\store'$-
$\NIMO$ guarantees ideal reconstruction of the real-world corrupt views. 
 However, there is a subtlety here in that the adversary
must be able to tractably reconstruct this knowledge. This is not
generally true, in particular, if $\idealf$ is a 1-way
function. Define:
\begin{definition}[Ideal Knowledge]
  Given a functionality $\idealf$ and output value $v$, the associated
  \emph{ideal knowledge}, denoted $\ik(\idealf,v)$ is
  $
  \{ \store\ |\ \idealf(\store) = v \}
  $.
  We say that $\idealf$ is \emph{pre-imageable} iff $\ik(\idealf, v)$ for all
  $v$ can be computed tractably.
\end{definition}
Now we can show our main result, that noninterference modulo output,
together with the assumption that $\idealf$ is pre-imageable, implies
passive security of a protocol. The proof is provided in Appendix
\ref{section-proofs}, and follows by essentially the same argument as
for the analogous property of ``perfect passive NI security'' in
\cite{8429300}.  
\begin{restatable}{theorem}{nimosecure}
  \label{theorem-nimo}
  Assume given pre-imageable $\idealf$ and a protocol $\prog$ that
  correctly implements $\idealf$.  If $\NIMO(\prog)$
  then $\prog$ is passive secure.
\end{restatable}
As for noninterference, we can show that $\NIMO$ is equivalent
to saying that a given program does not change the probability
of honest secrets conditioned on corrupt secrets and views
(including the output view), as compared to conditioning on
just the corrupt secrets and input. In addition to providing
insights about $\NIMO$, this form is convenient for
verification techniques. 
\begin{lemma}
  \label{lemma-nimo}
  $\NIMO(\prog)$ iff for all $H$ and $C$ with $|C| \le |H|$ and
  $\store \in \mems(S \cup V)$:
  $$\condd{\progd(\prog)}{S_H}{\store_{(S \cup \{\outv\})_C}} =
  \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}} $$
  where  $\iov(\prog) = (S,V)$.
\end{lemma}
